A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-06 kobakian-cook.Rmd
Rearrange figures and replace references

```{r setup}
library(knitcitations)
library(kableExtra)
library(RefManageR)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")

library(sp)
library(sf)
library(sugarbag)
# remotes::install_github("wfmackey/absmapsdata")
library(absmapsdata)
library(tidyverse)
library(cartogram)
library(ggthemes)
library(cowplot)

@@

In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC]. The creation of approximately equal population areas results in dramatically different square meterage of the administrative geographic units, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI]. This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:melanoma-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of melanoma cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Hobart stand out as more orange and red. To create this choropleth map the SA2 polygons for 2011 from the Australian Bureau of Statistics.

@@

```{=tex}
\begin{figure}[h]
\centering
\includegraphics[width=14cm]{figs/aus_melanoma_p.png}
\caption{\label{fig:melanoma-geo}A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Melanoma for males, they range from much lower than average (blue) to much higher than average (red)}
\end{figure}
```
Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing the population in each administrative area [@TAAM]. This acknowledges that the number of residents can be different but recognises that each area, or person within it is equally important.

@@

```{r thyroiddata}
states <- absmapsdata::state2011 %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  rmapshaper::ms_simplify(keep = 0.25, keep_shapes = TRUE)

sa2 <- absmapsdata::sa22011 %>% 
  filter(state_name_2011 == "Tasmania") %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!state_name_2011 == "Other Territories")

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf() 
```

```{r choro, eval=FALSE}
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(fill = "lightgrey", data = states %>% filter(state_name_2011=="Tasmania")) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +invthm
```

```{r fullhexmap}
## Create centroids set
centroids <- absmapsdata::sa22011 %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(state_name_2011 == "Tasmania") %>% 
 # filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  create_centroids(., "sa2_name_2011")

@@

fort_aus <- sa2thyroid_ERP %>%
 # filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  fortify_sfc()

fort_state <- states %>% 
  filter(state_name_2011 == "Tasmania") %>% 
  fortify_sfc() %>% 
  filter(long < 159)

## Make a plot
library(ggplot2)
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat, group = interaction(state_name_2011, polygon)),
               fill = "darkgrey",  data = fort_state) +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
#aus_hexmap
```

```{r alternatives}
sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112)
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

@@

cont <- sa2thyroid_ERP %>% 
  mutate(Population = Population + 1) %>% 
  cartogram_cont(., weight = "Population", itermax = 20) %>%
  st_as_sf()

save(cont, file = "data/cont.rda")
load("data/cont.rda")

aus_ggcont <- ggplot(cont) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
#aus_ggcont

@@

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  invthm + guides(fill=FALSE)
#tas_ggncont

# Hobart
hobart <- ggplot(ncont %>% filter(gcc_name_2011 == "Greater Hobart")) + 
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP %>% filter(gcc_name_2011 == "Greater Hobart")) +
  geom_sf(aes(fill = SIR, size = 0.001), colour = "black") + 
  scale_fill_identity() + 
  scale_size_identity() +
  invthm + guides(fill=FALSE) + 
  theme(plot.background = element_rect(fill = "transparent", colour = NA)) + 
  theme(panel.background = element_rect(fill = "transparent", colour = "white"))

aus_ggncont <- ggdraw(xlim = c(0, 1), ylim = c(0, 2)) +
  draw_plot(tas_ggncont, width = 1, height = 2) +
  draw_plot(hobart, x = 0, width = 0.4, height = 0.4)
# aus_ggncont

#ggsave(filename = "figs/aus_ggncont.png", plot = aus_ggncont,
#       device = "png", bg = "transparent", dpi = 300,  width = 7, height = 6)

@@

###############################################################################
# Non - Contiguous Dorling Cartograms

dorl <- sa2thyroid_ERP %>%
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "darkgrey",  data = states %>% st_transform(., 3112) %>% filter(state_name_2011 == "Tasmania")) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  invthm + guides(fill = FALSE)
# aus_ggdorl

@@

The choropleth map that uses the geographic shapes of areas can be seen in figure \ref{fig:melanoma-geo}. Figure \ref{fig:melanoma-hex} displays the Australian areas with equally sized hexagons. This display allows the small and densely populated inner city areas to be easily contrasted with their neighbouring areas.

>>>>

```{r setup}
library(knitcitations)
library(kableExtra)
library(sp)
library(sf)
library(sugarbag)
library(tidyverse)
library(cartogram)
library(ggthemes)
library(cowplot)

# remotes::install_github("wfmackey/absmapsdata")
library(absmapsdata)

library(RefManageR)
options("citation_format" = "pandoc")
BibOptions(check.entries = FALSE, style = "markdown")

@@

In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC]. The creation of approximately equal population areas results in dramatically different square meterage of the administrative geographic units, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI]. This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:thyroid-geo} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of melanoma cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of light blue areas across the rural communities in all states. The SA2s around Hobart stand out as more orange and red. To create this choropleth map the SA2 polygons for 2011 from the Australian Bureau of Statistics.

```{r thyroiddata}
sa2 <- absmapsdata::sa22011 %>% 
  filter(!st_is_empty(geometry)) %>% 
  filter(!state_name_2011 == "Other Territories") %>% 
  filter(!sa2_name_2011 == "Lord Howe Island")

sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  #filter(!is.na(Population)) %>% 
  filter(!sa2_name_2011 == "Lord Howe Island") %>% 
  mutate(SIR = map_chr(p50, aus_colours)) %>% 
  st_as_sf() 
```

```{r choro, fig.cap = "A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), size = 0.1) + 
  scale_fill_identity() + invthm
```

@@

Alternative mapping methods allow increased understanding of the spatial distribution of a variable across the population, by fairly representing the population in each administrative area [@TAAM]. This acknowledges that the number of residents can be different but recognises that each area, or person within it is equally important.

@@

```{r fullhexmap, fig.cap = "A choropleth hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
## Create centroids set
centroids <- sa2 %>% 
  create_centroids(., "sa2_name_2011")

@@

fort_aus <- sa2thyroid_ERP %>%
  fortify_sfc()

## Make a plot
aus_hexmap <- ggplot() +
  geom_polygon(aes(x = long,  y = lat,  group = interaction(sa2_name_2011, polygon)),
               fill = "black",  colour = "darkgrey",  size = 0.1, data = fort_aus) +
  geom_polygon(aes(x = long, y = lat, group = hex_id, fill = SIR), data = fort_hex) +
  scale_fill_identity() +
  invthm + coord_equal()
aus_hexmap
```

```{r alternatives}
# Project data 
sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112)
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

@@

if (!file.exists("data/ncont.rda")) {
cont <- sa2thyroid_ERP %>% 
  mutate(Population = Population + 1) %>% 
  cartogram_cont(., weight = "Population", itermax = 20) %>%
  st_as_sf()

save(cont, file = "data/cont.rda") }

load("data/cont.rda")

aus_ggcont <- ggplot(cont) +

  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + guides(fill=FALSE)
#aus_ggcont

@@

aus_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill=FALSE)
#aus_ggncont

@@

# Non - Contiguous Dorling Cartograms

dorl <- sa2thyroid_ERP %>%
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
aus_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "darkgrey",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  invthm + guides(fill = FALSE)
# aus_ggdorl

@@

The choropleth map that uses the geographic shapes of areas can be seen in figure \ref{fig:thyroid-geo}. Figure \ref{fig:melanoma-hex} displays the Australian areas with equally sized hexagons. This display allows the small and densely populated inner city areas to be easily contrasted with their neighbouring areas.
